Stuff related to tensor products of modal algebras

### 2020-06-08 tensor-product-stuff/Text.tex
update tensor stuff

  $\langle u_1, v_1 \rangle R^j_v \langle u_2, v_2 \rangle$ iff $u_1 = u_2$ and $u_1 R_2^j u_2$ for $1 \leq j \leq m$.
\end{center}
\end{defin}

@@

\begin{enumerate}
  \item commutativity 1
  \item commutativity 2
  \item confluence
\end{enumerate}

>>>>

  $\langle u_1, v_1 \rangle R^j_v \langle u_2, v_2 \rangle$ iff $u_1 = u_2$ and $u_1 R_2^j u_2$ for $1 \leq j \leq m$.
\end{center}
\end{defin}
This operation on Kripke frames commutes with disjoint unions, $p$-morphic images, and generated subframes.

@@

\begin{enumerate}
  \item (left commutativity) $\forall x, y, z \in W \: (x R_j y \: \& \: y R_i z \Rightarrow \exists u \in W \: (x R_i u \: \& \: u R_j z))$
  \item (right commutativity) $\forall x, y, z \in W \: (x R_i y \: \& \: y R_j z \Rightarrow \exists u \in W \: (x R_j u \: \& \: u R_i z))$
  \item (Confluence) $\forall x, y, z \in W (x R_j y \: \& \: x R_i z \Rightarrow \exists u \in W (y R_i u \: \& \: z R_j u))$
\end{enumerate}

The properties are expressed as modal formulas as well:
\begin{enumerate}
  \item ${\bf comm}^{\bf l}_{ij} = \Diamond_j \Diamond_i p \to \Diamond_i \Diamond_j p$
  \item ${\bf comm}^{\bf r}_{ij} = \Diamond_i \Diamond_j p \to \Diamond_j \Diamond_i p$
  \item ${\bf cr}_{ij} = \Diamond_i \Box_j p \to \Box_i \Diamond_j p$
\end{enumerate}

\begin{defin} Given unimodal modals logics $\mathcal{L}_1, \dots, \mathcal{L}_n$, the commutator

  \begin{center}
    $[\mathcal{L}_1, \dots, \mathcal{L}_n]$
  \end{center}
  is the smallest $n$-modal logic containing $\mathcal{L}_i$ and axioms ${\bf comm}^{\bf l}_{ij}$, ${\bf comm}^{\bf r}_{ij}$, and ${\bf cr}_{ij}$ for $i, j \in \{ 1, \dots, n\}$ with $i \neq j$.
\end{defin}

Since ${\bf comm}^{\bf l}_{ij}$, ${\bf comm}^{\bf r}_{ij}$, and ${\bf comm}^{\bf r}_{ij}$ are Salqvist formulas, one has

\begin{prop}
  If $\mathcal{L}_1, \dots, \mathcal{L}_n$ are canonical, then $[\mathcal{L}_1, \dots, \mathcal{L}_n]$ is canonical, and thus, elementary and Kripke complete.
\end{prop}

Moreover, one has

\begin{prop}
  $[\mathcal{L}_1, \dots, \mathcal{L}_n] \subseteq \mathcal{L}_1 \times \dots \times \mathcal{L}_n$.
\end{prop}

If the converse inclusion holds, then $\mathcal{L}_1, \dots, \mathcal{L}_n$ are \emph{product-mathcing}, see \cite{gabbay1998products}. The examples of product-matching logics are Horn axiomatisable ones \cite{kurucz2003many}. In particular, the following equality holds for Kripke complete and Horn axiomatisable logics $\mathcal{L}_1, \mathcal{L}_2, \mathcal{L}_3$:

\begin{center}
  $\mathcal{L}_1 \times \mathcal{L}_2 \times \mathcal{L}_3 = (\mathcal{L}_1 \times \mathcal{L}_2) \times \mathcal{L}_3 = \mathcal{L}_1 \times (\mathcal{L}_2 \times \mathcal{L}_3)$
\end{center}

Generally, this is an open question whether the product of modal logics is associative.
